---
title: "Reproducible Research: Peer Assessment 1"

---

```{r setwd, echo=FALSE}
rm(list=ls());gc()
setwd('C:/Users/cteitelbaum/Dropbox/reproducibleResearch')
```

##Loading and preprocessing the data

```{r load and examine data}
activity=read.csv('activity.csv')

names(activity)
summary(activity)
activity$date=as.Date(activity$date)
```

##What is mean total number of steps taken per day?

```{r histogram}
par(mfrow=c(1,1),mar=c(5,4,2,2))
stepsPerDay=tapply(activity$steps,activity$date,sum)
stepsPerDay=stepsPerDay[which(!is.na(stepsPerDay))]

hist(stepsPerDay,col=rgb(.5,.5,.5),breaks=20,
     xlab='Number of steps',main='Number of steps taken per day')
```

The mean number of steps taken each day is ```r mean(stepsPerDay)```.  
The median number of steps taken each day is ```r median(stepsPerDay)```.

##What is the average daily activity pattern?

```{r daily activity pattern}

stepsPerInterval=tapply(activity$steps,activity$interval,
                        function(steps) mean(na.omit(steps)))
intervals=unique(activity$interval)

plot(stepsPerInterval~intervals,main='Daily Activity Time Series',
     xlab='Interval',ylab='Mean number of steps',type='l')

```


##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max steps}
maxSteps=as.integer(which(stepsPerInterval==max(stepsPerInterval)))
```

Interval ```r maxSteps``` contains the maximum number of steps.

##Imputting missing values
```{r count missing values}
dataNA=which(is.na(activity$steps))
```
There are ```r length(dataNA)``` missing values in the data. I will replace these values
with the mean value for that interval.

```{r replace missing values}
dataNAIntervalID=activity[dataNA,'interval']

activityNoNA=activity
activityNoNA[dataNA,'steps']=stepsPerInterval[as.character(dataNAIntervalID)]
```


When missing values are replaced in this way, the distribution of steps taken
per day looks like this:
```{r histogram no missing data}

stepsPerDayNoNA=tapply(activityNoNA$steps,activityNoNA$date,sum)

hist(stepsPerDayNoNA,col=rgb(.5,.5,.5),breaks=20,
     xlab='Number of steps',main='Number of steps taken per day')

```

The mean number of steps taken per day is now ```r mean(stepsPerDayNoNA)```.  
The median number of steps taken per day is now ```r median(stepsPerDayNoNA)```.

On average, imputting the missing data values increases the number of steps per day
by ```r (mean(stepsPerDayNoNA)-mean(stepsPerDay))/mean(stepsPerDay)```%.  


##Are there differences in activity patterns between weekdays and weekends?

```{r weekends}

activityNoNA$dayName=weekdays(activityNoNA$date)
activityNoNA$weekend='weekday'
activityNoNA[which(is.element(activityNoNA$dayName,c('Saturday','Sunday'))),
             'weekend']='weekend'

par(mfrow=c(2,1),mar=c(1,3,3,3),oma=c(3,3,3,0))

temp=activityNoNA[which(activityNoNA$weekend=='weekday'),]
    
stepsPerInterval=tapply(temp$steps,temp$interval,
                            function(steps) mean(na.omit(steps)))
intervals=unique(temp$interval)
    
plot(stepsPerInterval~intervals,ylim=c(0,250),
         type='l',xlab='',ylab='',axes=F)
box()
axis(4)
mtext('Weekday')

par(mar=c(3,3,1,3))
temp=activityNoNA[which(activityNoNA$weekend=='weekend'),]
    
stepsPerInterval=tapply(temp$steps,temp$interval,
                            function(steps) mean(na.omit(steps)))
intervals=unique(temp$interval)
    
plot(stepsPerInterval~intervals,ylim=c(0,250),
         type='l',xlab='',ylab='',axes=F)
box()
axis(1);axis(2)
mtext('Weekend')

mtext('Daily Activity Time Series',outer=T,side=3,font=2)
mtext('Interval',outer=T,side=1)
mtext('Mean Number of Steps',outer=T,side=2)
```


